---
title: "R_reproducibility"
author: "Nhung"
date: "2025-04-01"
output: html_document
---

```{r setup, include = FALSE}

# required packages for reproducibility in R
library(here)
library(renv)
library(rticles)
library(devtools)
library(roxygen2)
library(gitcreds)

if (!require("devtools")) install.packages("devtools")
devtools::install_github("benmarwick/rrtools")

library(rrtools) # not yet installed
library(tinytex)
library(usethis)


# test the configuration for git/github/Rstudio
usethis::create_github_token()  # create a token
gitcreds::gitcreds_set()  # replace the token
usethis::gh_token_help() 
usethis::git_sitrep() # check the configuration of git/github/rstudio

credentials::set_github_pat() # for check

```
## Reproducibility (for research findings)

*Reproducibility means obtaining the same result by someone else given the same data and code.*

Robustness (durableness) means obtaining the same result from the same data using a different method. Exp: using different stats method on the same data to have the same conclusion on a hypothesis.

Replicability (of a research finding on different datasets) means obtaining the same result from an independent dataset using the same method. Exp: using the same bioinformatic pipeline on different datasets to achieve a same cataloge of mutations for Mtb.

Note for "here" package to make portable paths: to create absolute paths which runs the codes from paths that are relative to the root directory (a way to declare where the root directory is), using in 2 situations: 

(1) to fix the inconsistency of root directory during knit as knitting will automatically define the root working directory being the location of the knitted file, so here:here() will help to redefine the root directory by providing its relative path; 

(2) working with a script not inside the current project, so here::i_am() function will help to declare the relative path of the script to the its root project.

```{r here package}
# use here:here(.../...) to define the root directory by going back and searching the Rproj

# use here:i_am(.../...) before the here:here(.../...) to define the relative path of the root directory of the project


```

```{r compendium}
# first create a git-managed repo, then use the following codes to create a research compendium for the project

rrtools::use_compendium(simple = FALSE)  # create the skeleton
rrtools::add_dependencies_to_description()  # add the description
usethis::use_mit_license(copyright_holder = "Nhung")  # add licence
rrtools::use_readme_qmd()  # add the readme.Rmd
rrtools::use_analysis() # create the analysis folder, choose to make the file being tracked by git or not

# complex R code should be written as functions, make it as a package and documenting code,  put in R sub folder (codes can be written in code chunks, scripts or functions)

```


```{r renv }
# to create a project-specific R environment with its own library

renv::init() # create the scaffold of the project, use once at the beginning 
renv::status() # use many time when trying to install new packages in the project
renv::snapshot() # to update the packages into the lockfile

# renv for reproducibility: first need to download a reproducible project (just need to share the lockfile and activate.R file)

renv::restore()

```


```{r docker }

```
```{r useful functions}
usethis::create_from_github()  # clone the repo
renv::install("tidyverse") # install a package using renv

## {targets} pipeline

# step 1: turn a bunch of relevant codes into a function (such as a function to read and clean data, a function to create a bar plot, a function to create a scatter plot), then document the functions using {roxygen} (all the functions should be saved into a helper_function.R file and put into the R folder; to use them just need to source this script containing all the functions)

# step 2: turn the main script into a {targets} pipeline, using the package "targets". The pipeline is saved in a _targets.R file in the main directory (use the command below for targets configuration, setting up a custom directory)

targets::tar_config_set(script = "analysis/_targets.R", store = "analysis/_targets")
targets::tar_make()
targets::tar_read()

# using {assertr} for data checking

```





```{r cars, echo = FALSE}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
